---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r loadData}
library(data.table)
library(ggplot2)
library(dplyr)
```

## Loading and preprocessing the data
```{r}
unzip('activity.zip')
activity<-data.table(read.csv(file='activity.csv', header = T))
activity[, date:=as.Date(date, format='%Y-%m-%d')]
head(activity)
```


## What is mean total number of steps taken per day?
```{r}
TotalNumberOfSteps<-activity[, list(total_steps=sum(steps, na.rm = T)), by='date']
hist(TotalNumberOfSteps$total_steps, 
     main="Step Breakdown by Day" ,
     xlab='Total Number of Steps',
     col="green",
     breaks = 20)

meanSteps<-mean(TotalNumberOfSteps$total_steps, na.rm = T)
medianSteps<-median(TotalNumberOfSteps$total_steps, na.rm = T)
paste('The mean number of steps is', meanSteps)
paste('The median number of steps is', medianSteps)
```

## What is the average daily activity pattern?
```{r}
average_steps<-activity[!is.na(steps),list(steps=mean(steps)), by='interval']
average_steps[ ,steps:=as.numeric(steps)]
p1<-ggplot(data=average_steps, aes(x=interval, y=steps)) + 
        geom_line(color='green')+
        theme_bw()+
        ggtitle('Mean Steps by Interval')
print(p1)

max_step<-max(average_steps$steps)
max_interval<-average_steps[steps==max_step,]
max_interval<-max_interval$interval
paste(max_interval, 'is the interval that contains the max number of steps which is', max_step)
```


## Imputing missing values
### Empty values in this data set have been replaced by the meanInterval
```{r}
empty_row<-nrow(activity[is.na(steps),])
meanInterval<-mean(activity$steps, na.rm=T)
data_fill<-activity[, steps:=ifelse(is.na(steps), meanInterval, steps)]
updateNumberSteps<-data_fill[, list(total_steps=sum(steps, na.rm = T)), by='date']
updateMeanStep<-mean(updateNumberSteps$total_steps)
updatemediaStep<-median(updateNumberSteps$total_steps)

hist(updateNumberSteps$total_steps, 
     main="Step Breakdown by Day" ,
     xlab='Total Number of Steps',
     col="green",
     breaks = 20)

```
##With the empty values replaced, the updated mean and median values are
```{r}
paste(updateMeanStep, 'and', updatemediaStep,'respectively')
```
###This is a slight elevation on both values

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekdayOrWeekend<-copy(data_fill)
weekdayOrWeekend[,day:=weekdays(date)]
daytype<-c('Saturday','Sunday')
weekdayOrWeekend[,dayType:=as.factor(ifelse(day %in% daytype, 'weekend','weekday'))]
weekdayOrWeekend<-weekdayOrWeekend[, list(mean_steps=mean(steps)), by=c('interval','dayType')]

p2<-ggplot(weekdayOrWeekend, aes(x=interval, y=mean_steps, color=dayType)) + 
        geom_line()+
        xlab("Interval") + 
        ylab("Mean Number of Steps") +
        ggtitle("Average Number of Steps by Weekend and Weekday")+
        facet_grid(dayType ~.)
print(p2)
```






